title: "covid 19"
author: "Taelynn Condoll"
date: "summer 2022"
output:
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May

### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well  

# Question 1 - Is dying independent of sex

$H_0$: There is no relation between dying and sex

$H_A$: There is a relation between dying and sex

Table of patients:

```{r}
# cross tabulation, patient state by gender
with(kRdata, table(Sex, result))
```

The table shows that death is related to sex due to only 25 females dying as opposed to 44 males dying.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(Sex, result) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_followed by  summarize(count=n())
basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percent plot
basicC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result))
basicC + geom_bar(stat="identity",position="dodge")
```

These bar graphs shoe that males are more likley to die then females . The first graph displays the frequency of death and the second shows the precent dead . males have a much higher chance of death.

## Numerical results

```{r}
table1 <- xtabs(~Sex + result, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The table shows men are significantly more likley to die then women.

## Inferential results

```{r}
table1 <- xtabs(~Sex + result, data=kRdata)
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p val is .0009 which is smaller then .05 so we reject the null hypothesis.

## Conclusions

This study shows men are much more likely to die then women. The p-value is significantly less then 0.05. Therefore we reject the null hypothesis which states that there is no relation between death and sex.
